Two subtle problems with over-representation analysis

### 2024-06-14 manuscript/manuscript.Rmd
resources

Secondly ORA results can vary dramatically depending on the number of genes selected in the
foreground, a mostly arbitrary choice.
These issues together with previously mentioned pitfalls such as lack of background
correction and methodological misreporting [@Timmons2015-ki;@Wijesooriya2022-fw], suggest
that ORA should be discouraged in most situations, and replaced by other methods such as FCS.
Where it makes sense to use ORA would be when FCS is not applicable, for example when
investigating the over-represented categories in gene expression modules defined by packages
like WGCNA (Weighted Gene Correlation Network Analysis)[@Langfelder2008-hu], or when
investigating over-represented functional categories among genes with variants associated
with a disease.

@@

* The code repository including template Dockerfile and R Markdown script are available on GitHub
(https://github.com/markziemann/enrichment_recipe).

* The example Docker image is available on DockerHub (https://hub.docker.com/r/mziemann/enrichment_recipe).

>>>>

The lack of enrichment scores hampers interpretation.

Lastly, ORA results can vary dramatically depending on the number of genes selected in the
foreground, a mostly arbitrary choice.
These issues together with previously mentioned pitfalls such as lack of background
correction and methodological misreporting [@Timmons2015-ki;@Wijesooriya2022-fw], suggest
that ORA should be discouraged in most situations, and replaced by other methods such as FCS.
Where it makes sense to use ORA would be when FCS is not applicable, for example when
investigating the over-represented categories in gene expression modules defined by packages
like WGCNA (Weighted Gene Correlation Network Analysis)[@Langfelder2008-hu], or when
investigating over-represented functional categories among genes with variants associated
with a disease.

@@

* The code repository including template Dockerfile and R Markdown script are available on
GitHub (https://github.com/markziemann/background).

* The example Docker image is available on DockerHub
(https://hub.docker.com/r/mziemann/background).
